---
title: "Lab 04: Analysis of Variance"
author: "Isaac Plotkin"
date: "2/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(broom)
```

```{r}
data(diamonds)

diamonds_5c <- subset(diamonds, carat == 0.5)

```
There are 1,258 observations in the new dataset.

```{r}
p1 <- ggplot(data = diamonds_5c, aes(x = cut)) +
  geom_bar() + 
  labs(title = "Diamond Cut Quality")

p1
```

Fair and Good cuts have the fewest number of observations.


```{r}
diamonds_new <- mutate(diamonds_5c, cut = fct_recode(cut, "FairGood" = "Fair", "FairGood" = "Good"))

diamonds_new
```
The variable cut got recoded correctly.


```{r}
pvc <- ggplot(data = diamonds_new, aes(x = cut, y = price)) +
  geom_boxplot() +
  labs(title = "Price vs. Cut")

pvc
```
```{r}
print(summary(diamonds_new$cut))
summary <- aggregate(price~cut, diamonds_new, function(x) c(mean = mean(x), sd = sd(x)))

summary[-1][[1]]
```

There appears to be a relationship between the cut and price for diamonds that are 0.5 carats because the box plot and summary statistics show
that the mean is increasing as the quality of the cut increases. 


# ANOVA

The assumptions for ANOVA are satisfied because the distribution of price appears normal within each cut category. 
All the observations are collected separately so independence is satisfied. Lastly constant variance is satisfied because all of the standard
deviations are close in size.


```{r}
model <- lm(price ~ cut, data = diamonds_new)
tidy(model, conf.int = TRUE) %>%
  kable(format = "markdown", digits = 3)

kable(anova(model), format="markdown", digits = 3)
```

Total variation = SST/DFT
```{r}
total_variation <- 150706506/1257
total_variation
```
















